---
title: "Kuo et al 2010 CgAp-1 Reanalysis"
author: Bin He
output: html_notebook
---

**Goal**

Reanalyze Yap1 ChIP in _C. glabrata_ in order to identify the genomic regions bound by the TF.

**Reference**
1. Kuo D, Licon K, Bandyopadhyay S, Chuang R, Luo C, Catalana J, et al. Coevolution within a transcriptional network by compensatory trans and cis mutations. Genome Res. 2010 Dec 1;20(12):1672–8. 

**Data**
Three replicates for CgAp-1 ChIP. For each replicate, the IP enriched DNA is labeled with Cy5 (and the corresponding unenriched (input) DNA is labeled with Cy3. The mixture of the two was hybridized to a custom Agilent microarray, scanned using a Genepix 4000A scanner and analyzed by GenePix 6.0, producing a two-color array dataset in the gpr format.

## Setup the environment

```{r load_libraries}
require(limma)
require(ggplot2)
require(Biobase)
require(GEOquery)
```

## Data preparation

There are three files in total, corresponding to the three replicates, with the names of GSM397447,48,49

```{r load_data}
# load target data frame, containing the sample information and file names
targets <- readTargets("../data/raw-chip-gpr/yap1-chip-sample.txt",row.names = "Name")
# add relative path
targets$FileName <- paste("../data/raw-chip-gpr/",targets$FileName,sep="")
# load raw data
RG <- read.maimages(targets, source="genepix")
# load platform (microarray) design
gpl8477 <- getGEO(filename = "../data/raw-chip-gpr/GPL8477_family.soft.gz")
```

```{r data_QC}
# MA plot
plotMD(RG)
# Boxplot for the background intensities
boxplot(data.frame(log2(RG$Gb)),main="Green background")
boxplot(data.frame(log2(RG$Rb)),main="Red background")
# array 2 seems to have unusually high background, examine its image
imageplot(log2(RG$Gb[,2]),RG$printer, main = "array 2, Green")
imageplot(log2(RG$Rb[,2]),RG$printer, main = "array 2, Red")
# compare them to array 1 and 3
imageplot(log2(RG$Gb[,1]),RG$printer, main = "array 1, Green")
imageplot(log2(RG$Rb[,1]),RG$printer, main = "array 1, Red")
imageplot(log2(RG$Gb[,3]),RG$printer, main = "array 3, Green")
imageplot(log2(RG$Rb[,3]),RG$printer, main = "array 3, Red")
```

```{r pre_processing}
RGb1 <- backgroundCorrect(RG, method="subtract")
RGb2 <- backgroundCorrect(RG, method="normexp", offset=50)
plotMD(RGb1, column=2)
plotMD(RGb2, column=2)
# the normexp method was recommended by LIMMA manual. within array normalization is then applied on the background corrected dataset
MA <- normalizeWithinArrays(RGb2)
```