---
title: "Code_Figure_2d"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#allow fread function and data.table structure
library(data.table)
#allow heatmap functions
library(gplots)
#allow merging data tables
library(dplyr)

#Reference data/output directories
data_dir <- "../data"
analysis_dir <- "../analysis"
output_dir <- "../output"

#Load objects from earlier analysis
load(file = "filtered_data.RData")
load(file = "filtered_data_final.RData")
load(file = "filtered_data_mena.RData")
load(file = "filtered_data_mena_final.RData")
load(file = "genenames.RData")
load(file = "genenames_mena.RData")

```

## Goal

The goal of this analysis is to use the normalized data set provided to attempt to reproduce the results as well as identify any abiguity in the methods described in the original study.

## Methods Used

The data was imported from the given tab-delmited file provided in the paper and exported as a .csv file for use in R: "normalized_data_oxidative_stress_Candida_glabrata.txt.csv". The analysis was done by formatting the data in the .csv. 

##Legend - Based on .sdrf file & Heatmap 2b

| Column | Sample 1 | Sample 2 |
| --------- | ----------- | ----------- |
| 1  | WT Glucose starvation | WT unstressed |
| 2  | WT Menadione stress | WT unstressed |
| 3  | yap1$\Delta$skn7$\Delta$ Menadione stress | yap1$\Delta$skn7$\Delta$ unstressed |
| 4  | WT H~2~O~2~ stressed | WT unstressed |
| 5  | yap1$\Delta$skn7$\Delta$ H~2~O~2~ stressed | yap1$\Delta$skn7$\Delta$ unstressed |



```{r format csv, echo=FALSE}
#Turn off warnings for readablility
options(warn = -1)

#Combine data sets from H2O2 and Mena into one
filtered_data <- inner_join(filtered_data, filtered_data_mena, by="Genes")

# Only keep rows where atleast one column has >2x induction/repression, 1 is used here since data was log-base 2 transformed, also account for CV<1 to remove highly variant genes
filtered_data <- filtered_data[(abs(filtered_data$`Glucose Data`) > 2) & ((abs(filtered_data$`1.x`) > 2 & abs(filtered_data$`1 CV.x`) < 1) | (abs(filtered_data$`1.x`) > 2 & abs(filtered_data$`1 CV.x`) < 1)),] 

#save final list of gene names
new_genenames <- filtered_data$Genes

filtered_data <- filtered_data[, c("Genes", "Glucose Data", "1.y", "2.y", "1.x", "4")]

#Remove any rows with NAs in them
filtered_data <- filtered_data[complete.cases(filtered_data), ]

#convert to matrix for easy use of heatmap
filtered_data <- data.matrix(filtered_data)

#Testing
#print(names(filtered_data))
#print(filtered_data)

```


```{r generate heatmaps and print, echo=FALSE}
#Turn off warnings for readablility
options(warn = -1)

# Replace the old distance function to account for NA values in dataset
dist_no_na <- function(mat) {
    edist <- dist(mat)
    edist[which(is.na(edist))] <- max(edist, na.rm=TRUE) * 1.1 
    return(edist)
}

#Assign color gradient for heatmap
colorway <- colorRampPalette(c("cyan", "black", "yellow"))(n = 500)

#Heatmap with legend
heatmap.2(filtered_data[,2:ncol(filtered_data)],
  labRow = genenames,
  main = "Fig. 2d", # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  margins =c(1,8),     # widens margins around plot
  col=colorway,       # use on color palette defined earlier
  dendrogram="none",     # only draw a row dendrogram
  Colv="NA",
  distfun=dist_no_na,   # needed for NA values to not interfere in heatmap
  cexCol = 1,
  key.xlab = "Induction Values (log2 transformed)"
)


```
