---
title: "Code_Figure_2a"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#allow fread function and data.table structure
library(data.table)
#allow heatmap functions
library(gplots)
#allow merging data tables
library(dplyr)

#Reference data/output directories
data_dir <- "../data"
analysis_dir <- "../analysis"
output_dir <- "../output"


```

## Goal

The goal of this analysis is to use the normalized data set provided to attempt to reproduce the results as well as identify any abiguity in the methods described in the original study.

## Methods Used

The data was imported from the given tab-delmited file provided in the paper and exported as a .csv file for use in R: "normalized_data_oxidative_stress_Candida_glabrata.txt.csv". The analysis was done by formatting the data in the .csv. 

Initially, all columns unrelated to the direct heatmap production were removed. Following this, only rows containing a 8x induction or repression were preserved. In the study this detail was left ambiguous. To instead use a threshold of 3x induction/repression, a value of 1.584 would be used below, resulting in 335 genes being found instead of 29 in this analysis and the 38 in the paper.

Once the genes were identified, the values were put into a matrix. To account for NA values in the data provided, a modified distance function was used so clustering would be an option in the heatmap. The modified function replaces NA values in the dataset with 1.1 times the max euclidean distance allowable. The benefit of this is to allow the clustering function based on euclidean distances to work while also letting the NA values get split from the rest of the data quickly. The result of this produces 3 distinct clusters of genes which supports one of the goals of the study. The clustering done in the paper itself wasn't explicity stated. Lastly, a dendogram was provided to further show relationships between induced/repressed genes.


##Legend - Based on .sdrf file & Heatmap 2a

| Column | Sample 1 | Sample 2 |
| --------- | ----------- | ----------- |
| 1  | WT stressed | WT unstressed |
| 2  | WT stressed | skn7$\Delta$ stressed |
| 3  | WT stressed | yap1$\Delta$ stressed |
| 4  | yap1skn7$\Delta$ stressed | yap1skn7$\Delta$ unstressed |
| 5  | WT Glucose starvation | WT unstressed |


```{r read data, echo=FALSE}

#read in data from csv for non-glucose data
fread("../data/normalized_data_oxidative_stress_Candida_glabrata.txt.csv",  na.strings=getOption("datatable.na.strings","null")) -> unfiltered_data

#read in data from csv for glucose data
fread("../data/normalized_Cg_glucose_data.csv",  na.strings=getOption("datatable.na.strings","null")) -> unfiltered_data_glucose

```

```{r format csv, echo=FALSE}
#Turn off warnings for readablility
options(warn = -1)

#Remove header columns and rows to get raw data
unfiltered_data[-1,] -> filtered_data
unfiltered_data_glucose[-1,] -> filtered_data_glucose

#Remove SD and CV columns since not needed for heatmap below - kept CV rows
filtered_data <- filtered_data[,-c(3,6,8,9,10,12,14,15,16,18)]

#For glucose data, only keep the single Mean value column and names
filtered_data_glucose <- filtered_data_glucose[,c(1,5)]

#This line is just used to make the next step easier, not necessary; see note below about bug
names(filtered_data) <- c("Genes", "4", "4 CV", "3", "3 CV", "2", "2 CV", "1", "1 CV")

#extra column created to work around bug mentioned below
filtered_data$dup <- filtered_data$`1`

#Rearrange columns to match legend
filtered_data <- filtered_data[,c(1,8,9,6,7,4,5,2,3,10)]

#Rename glucose columns so you can merge
names(filtered_data_glucose) <- c("Genes", "Glucose Data")

#Merge the two data tables
filtered_data <- inner_join(filtered_data, filtered_data_glucose, by="Genes")

#convert joined data frame back to data table
filtered_data <- as.data.table(filtered_data)

#Rearrange columns to match legend and account for glucose column due to bug
filtered_data <- filtered_data[,c(1,8,9,6,7,4,5,2,3,11,10)]

#converts values checked by str() from factors to numerics for heatmap 
#There is a bug here where the first column (2 in this case) becomes all NA - col 'dup' from above is used as an extra column that gets the values from the column to the left of it, while the second column gets nullified so that all the important columns are ok - second column will now get removed in next step and everything will be renamed accordingly to the values in the .csv
filtered_data[, 2:ncol(filtered_data) := lapply(.SD, as.numeric)]

#remove nullified column
filtered_data <- filtered_data[, -2]

#Get all the correct column names in order
names(filtered_data) <- c("Genes", "1", "1 CV", "2", "2 CV", "3", "3 CV", "4", "4 CV", "Glucose Data")

#save filtered_data file for use in other .Rmd files
save("filtered_data", file = "2a.RData")

# Only keep rows where atleast one column has >3x induction/repression, 1.59 is used here since data was log-base 2 transformed, also account for CV<1 to remove highly variant genes
filtered_data <- filtered_data[((abs(filtered_data$`1`) > 1.59) & filtered_data$`1 CV`<1) | ((abs(filtered_data$`2`) > 1.59) & filtered_data$`2 CV`<1) | ((abs(filtered_data$`3`) > 1.59) & filtered_data$`3 CV`<1) | ((abs(filtered_data$`4`) > 1.59) & filtered_data$`4 CV`<1),] 

#Remove CV columns for heatmap analysis
filtered_data <- filtered_data[,-c(3,5,7,9)]

#Save the list of names for later use in labeling
genenames <- filtered_data$`Genes`

#Remove any rows with NAs in them
filtered_data <- filtered_data[complete.cases(filtered_data), ]

#convert to matrix for easy use of heatmap
filtered_data <- data.matrix(filtered_data)

#Testing
#print(filtered_data)
#print(genenames)
#print(nrow(filtered_data))

```


```{r generate heatmaps and print, echo=FALSE}
#Turn off warnings for readablility
options(warn = -1)

# Replace the old distance function to account for NA values in dataset
dist_no_na <- function(mat) {
    edist <- dist(mat)
    edist[which(is.na(edist))] <- max(edist, na.rm=TRUE) * 1.1 
    return(edist)
}

#Assign color gradient for heatmap
colorway <- colorRampPalette(c("cyan", "black", "yellow"))(n = 500)

#Heatmap with legend 
fig2a <- heatmap.2(filtered_data[,2:ncol(filtered_data)],
  labRow = genenames,
  main = "Fig. 2a",     # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  margins =c(1,8),      # widens margins around plot
  col=colorway,         # use on color palette defined earlier
  dendrogram="none",    # only draw a row dendrogram
  Colv="NA",
  distfun=dist_no_na,   # needed for NA values to not interfere in heatmap
  cexCol = 1,
  key.xlab = "Induction Values (log2 transformed)"
)


```

