---
title: "Code_Figure_2b"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    theme: united
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#allow fread function and data.table structure
library(data.table)
#allow heatmap functions
library(gplots)

#Reference data/output directories
data_dir <- "../data"
analysis_dir <- "../analysis"
output_dir <- "../output"


```

## Goal

The goal of this analysis is to use the normalized data set provided to attempt to reproduce the results as well as identify any abiguity in the methods described in the original study.

## Methods Used

The data was imported from the given tab-delmited file provided in the paper and exported as a .csv file for use in R: "normalized_data_oxidative_stress_Candida_glabrata.txt.csv". The analysis was done by formatting the data in the .csv. 

##Legend - Based on .sdrf file

| Column | Sample 1 | Sample 2 |
| --------- | ----------- | ----------- |
| 1  | WT Menadione Stress | WT unstressed |
| 2  | yap1skn7 Menadione Stress | yap1skn7 Unstressed |

```{r read data, echo=FALSE}

#read in data from csv
fread("../data/normalized_data_oxidative_stress_Candida_glabrata.txt.csv",  na.strings=getOption("datatable.na.strings","null")) -> unfiltered_data

```

```{r format csv, echo=FALSE}
#Turn off warnings for readablility
options(warn = -1)

#Remove header columns and rows to get raw data
unfiltered_data[-1,] -> filtered_data

#Remove any rows with NAs in them
filtered_data <- filtered_data[complete.cases(filtered_data), ]

#Remove SD and CV columns since not needed for heatmap below - keep 8,10,14,16 since Mena columns
filtered_data <- filtered_data[,-c(2,3,4,5,6,7,9,11,12,13,15,17,18,19)]

#This line is just used to make the next step easier, not necessary; see note below about bug
names(filtered_data) <- c("Genes", "1", "1 CV", "2", "2 CV")

#extra column created to work around bug mentioned below
filtered_data$dup <- filtered_data$`1`

#converts values checked by str() from factors to numerics for heatmap 
#There is a bug here where the first column (2 in this case) becomes all NA - col 'dup' from above is used as an extra column that gets the values from the column to the left of it, while the second column gets nullified so that all the important columns are ok - second column will now get removed in next step and everything will be renamed accordingly to the values in the .csv
filtered_data[, 2:ncol(filtered_data) := lapply(.SD, as.numeric)]

#remove nullified column
filtered_data <- filtered_data[, -2]

#Reset names to correct values
names(filtered_data) <- c("Genes", "1", "1 CV", "2", "2 CV")

# Only keep rows where atleast one column has >3x induction/repression, 1.59 is used here since data was log-base 2 transformed, also account for CV<1 to remove highly variant genes
filtered_data <- filtered_data[((abs(filtered_data$`1`) > 1.59) & filtered_data$`1 CV`<1) | ((abs(filtered_data$`2`) > 1.59) & filtered_data$`2 CV`<1),]

#Remove CV columns for heatmap analysis
filtered_data <- filtered_data[,-c(3,5)]

#Save the list of names for later use in labeling
genenames <- filtered_data$`Genes`

#convert to matrix for easy use of heatmap
filtered_data <- data.matrix(filtered_data)

#Testing
#print(names(filtered_data))
print(filtered_data)

```

## Heatmap of Figure 2b

```{r generate heatmaps and print, echo=FALSE}
#Turn off warnings for readablility
options(warn = -1)

# Replace the old distance function to account for NA values in dataset
dist_no_na <- function(mat) {
    edist <- dist(mat)
    edist[which(is.na(edist))] <- max(edist, na.rm=TRUE) * 1.1 
    return(edist)
}

#Assign color gradient for heatmap
colorway <- colorRampPalette(c("cyan", "black", "yellow"))(n = 500)

#Heatmap with legend - add row names
#Note: I am not sure how to approach the NA data issue since it causes blank areas in my heatmap but doesn't seem to affect the data in the paper --- come back and look at this after finishing other parts
heatmap.2(filtered_data[,2:ncol(filtered_data)],
  labRow = genenames,
  main = "Fig. 2a", # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  margins =c(1,8),     # widens margins around plot
  col=colorway,       # use on color palette defined earlier
  dendrogram="none",     # only draw a row dendrogram
  Colv="NA",
  distfun=dist_no_na,   # needed for NA values to not interfere in heatmap
  cexCol = 1,
  key.xlab = "Induction Values (log2 transformed)"
)


```

